---
title: "The Importance of Breast Cancer Risk Awareness: A Thorough Analysis"
author: "Nicole Ayala"
date: "`r Sys.Date()`"
output: 
    prettydoc::html_pretty:
    toc: TRUE
    theme: HPSTR
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

### The Purpose of This Independent Project

 - To focus on some factors that correlate to the cause of breast cancer that then impact the possibility of the cancer recurring due to said factors. 
 - To investigate how these factors associate with each other. 

### Questions to Answer
 - How do the quantitative variables matter when it comes to determining breast cancer malignancy?
 - Does age have a huge affect when it comes to severity of malignancy and recurrence probability?
 - What particular section of the breast is most likely to be diagnosed with cancer with a large tumor?
 - Are middle aged individuals more common to undergo irradiation therapy compared to those 50 and older?


### Load the Libraries

```{r}
library(tidyverse)
library(here)
library(ggtext)
library(ggplot2)
library(kableExtra)
library(gdata)
library(ggrepel)
library(ggridges)
library(hrbrthemes)
library(beyonce)
library(dplyr)
library(tidyr)
library(gganimate)
```


### Read in the Data
```{r}
cancer <- read_csv(here("Independent Project", "Data", "breast-cancer.csv")) # the breast cancer data set
breastcancer <- read_csv(here("Independent Project", "Data","data.csv")) 
#View(breastcancer)
```

### Filter the Data
```{r}
hexcancer <- cancer %>% # piping cancer dataset to hexcancer 
  drop_na() %>% # remove any unnecessary na's
  filter(class == "recurrence-events") # only include recurrence event for class (those that had the cancer come back)
barcancer <- cancer %>% # piping cancer dataset to barcancer
  drop_na() # remove any unnecessary na's
heatcancer <- breastcancer %>% # piping cancer dataset to heatcancer
  #filter(diagnosis == "M") %>% # only include malignant tumor  cancer diagnoses
  filter(symmetry_mean > ".2106") # want to have percentages higher than 25% as those represent higher breast density fatty tissue rates of 3 and 4
```

### HeatMap w/ Hexagons: Average Radius vs Average Texture
```{r The Hex Plot, fig.asp= 1.0, fig.align='center', fig.dim = c(7, 8), fig.cap= "This figure shows the averages of texture sensitivity of the breast from the center point of the core tumor based on their average radii, organized by the diagnosis of a benign or malignant tumor.", echo=FALSE}
  ggplot(data = breastcancer,aes( # attaching my dataset
                x = texture_mean, # x axis
                y = radius_mean,  # y axis
                fill = diagnosis
                )) + 
         geom_hex(stat = "binhex", position = "identity") + # wanted to do a hex heatmap
  labs(title = "Correlation of the Two Frequencies,\n Average of Radius and Texture within Breast Cancer Patients", subtitle = "There are two groups of breast cancer patients,some with benign tumors and others with malignant tumors.", # title and subtitles  
       x="Average Texture", # x axis but detailed
       y="Average Radius", # y axis but detailed
       fill = "Diagnosis") + # this plot is organized by diagnosis
  scale_fill_manual(values = beyonce_palette(9)) + # the color palette all blue :)
  theme_classic() + # classic theme
  theme(plot.title = element_text(size = 15, hjust = 0.5, color ="darkseagreen4"), # title size and color
        plot.title.position = "plot", # title adjustments
        plot.subtitle = element_text(size =10,hjust = 0.5, lineheight = 1.1, color = "darkseagreen"), # subtitle size and color
        axis.title = element_text(size=15,color = "darkorchid4", hjust = 0.5), # title adjustments
        axis.text = element_text(size = 10,color = "darkorchid4",hjust = 0.5), # text adjustments
        panel.grid = element_line(linewidth = 0.35, linetype = "dotted", color = "darkorchid4"), # background gridlines
        plot.margin = unit(c(1.0, 1.0, 1.0, 1.0), "cm"), # margin adjustments
        plot.background = element_rect(color = "white", fill ="white")) # background color
        
        ggsave(here("Independent Project","Output","hexplot.png")) # saving the plot!!
```
## **How do the quantitative variables matter when it comes to determining breast cancer malignancy?**

According to the figure above, these quantitative variables, specifically here average of both breast radii and texture demonstrate that individuals who tend to have higher radii(the average total of distance from the areola to the primary site of where the cancer was discovered) is more likely to have a higher texture (measures standard deviation of the gray scale values of the imaging done) so if there were some abnormal benign or malignant micro/macrocalcifications, (excessive white fatty breast tissue) it would most likely be malignant if both average texture and mean were above the average. In a normal mammogram, normal fatty tissue would be all dark with little streaks of dense white fatty tissue, but here it's not the usual case.

![Breast Calcification](https://www.verywellhealth.com/thmb/iGU4zAtk1LcjDwogSj2AZRZEucc=/750x0/filters:no_upscale():max_bytes(150000):strip_icc():format(webp)/calcifications-56a0b6835f9b58eba4b30bc3.jpeg){ width="200" height="200" }


![Cancerous Breast Tumor](https://www.verywellhealth.com/thmb/3Zchi6sRk_hAsU5U_PO9ivguhXI=/750x0/filters:no_upscale():max_bytes(150000):strip_icc():format(webp)/lateral-mammogram-of-female-breast-with-tumor-92263689-813095ee469b45eabfc9f5f4747758ed.jpg){ width="200" height="200" }

### Bubble Plot : Breast Quad vs Tumor Size
```{r The Bubble Quad Plot, fig.asp= 1.0, fig.align='center', fig.cap= "This bubble plot demonstrates how older breast cancer patients (aged 50+) tend to commonly get tumors on the bottom left breast with a malignancy degree of 3.", echo=FALSE}
ggplot(data = subset(cancer,!is.na(`breast-quad`)), # data is breastcancer
       aes(x = `tumor-size`, # x axis
           y = `breast-quad`,  # y axis
              size = `deg-malig`, # size 
              frame = age,col = age)) + # frame and column are both age
  geom_point(shape = 17) + # shape 17 is a triangle and using geom point :)
  geom_label_repel(aes(label = `deg-malig`, # using geom_repel to make labels for malignant rates
                  box.padding   = 0.35, # box surrounding the malignant rate
                  point.padding = 0.5, # size of number representing degree of malignancy
                  segment.color = 'black')) + # color of box
  scale_x_discrete(guide = guide_axis(n.dodge = 2)) + # color of the guide
  scale_color_manual(values = c("30-39" = "darkgreen", "40-49" = "firebrick3", "50-59" = "deepskyblue", "60-69" = "gold2", "70-79"="antiquewhite4")) + # used discrete value colors to color in the age differences
  theme_tinyhand() + # the theme
   labs(title = "The Impact Between Tumor Size and Breast Quadrant on \n Malignancy Severity at Various Ages: A Comprehensive Analysis", # title
       subtitle = "Most common areas in the breast that develop a tumor of a 2-3 malignancy \n is the center areola and the lower and upper quadrants of the right side.", # subtitle
    x = "Size of Tumor",  # x axis
    y = "Breast Quadrant",# y axis
    size = "Degree of Malignancy", # size
    frame = "Age", col = "Age") + # frame and column represent age
   transition_states( 
        age, # what are we animating by
        transition_length = 4, #The relative length of the transition.
        state_length = 3) + # The length of the pause between transitions
        ease_aes("linear") + # type of animation
  theme(plot.title = element_text(size = 14, hjust = 0.5, color ="antiquewhite4"), # title size and color
        plot.title.position = "plot", # title adjustments
        plot.subtitle = element_text(size =10,hjust = 0.5, lineheight = 1.1, color = "antiquewhite4"), # subtitle size and color
        axis.title = element_text(size=20,color = "darkgreen", hjust = 0.5), # title adjustments
        axis.text = element_text(size = 15,color = "darkgreen",hjust = 0.5), # text adjustments
        panel.grid = element_line(linewidth = 0.35, linetype = "dotted", color = "beige"), # background gridlines
        plot.margin = unit(c(1.0, 1.0, 1.0, 1.0), "cm"), # margin adjustments
        plot.background = element_rect(color = "white", fill ="white")) # background color
anim_save(here("Independent Project","Output","bubbleplotbreastquad.gif")) # saving the animation!
```
## **What particular section of the breast is most likely to be diagnosed with cancer with a large tumor?**

According to the data from Wisconsin, in their pool of patients, the quadrant of the breast that was most common when it came to diagnosing with a tumor without looking at age was the lower left quadrant. The range of malignancy ranged from 2 to 3 which means medium to severe (aggressive tumors).The ages that the patients were at the time of diagnosis with a concerning degree of malignancy was between the ages of 50-59 and 60-69. Those older than 70+ usually can not withstand having cancer and ultimately pass away sooner than those with a stronger immune system. 

_A suggestion for this would be to have an additional data column allocated to immuno-compromised patients to see the difference between those with a better immune system have a greater chance at surviving and conquering breast cancer._



###  Bubble Plot: Class vs Tumor Size
```{r The Bubble Class Plot, fig.asp= 1.0, fig.align='center', fig.cap= "In this figure,older breast cancer patients in the age range of 70-79 usually do not survive with severe malignant tumors, and for most of the individuals before that certain age, usually do not have a recurrence of the cancer coming back.", echo=FALSE}
ggplot(data = cancer, # data is breastcancer
       aes(x = `tumor-size`, # x axis
           y = `class`,  # y axis
              size = `deg-malig`, # size 
              frame = age,col = age)) + # frame and column are both age
  geom_point(shape = 20) + # shape 2 is a tiny circle and using geom point :)
  geom_label_repel(aes(label = `deg-malig`, # using geom_repel to make labels for malignant rates
                  box.padding   = 0.35, # box surrounding the malignant rate
                  point.padding = 0.5, # size of number representing degree of malignancy
                  segment.color = 'black')) + # color of box
  guides(x = guide_axis(angle = 90)) + # color of the guide
  scale_color_manual(values = c("30-39" = "plum", "40-49" = "pink", "50-59" = "lightskyblue", "60-69" = "bisque1", "70-79"="darkgray")) + # used discrete value colors to color in the age differences
  theme_tinyhand() + # the theme
   labs(title = "The Probability of Recurrence Events Ocurring \n
        When Tumors are Greater Than 20 mm.", # title
       subtitle = "Ages 50-59 and 60-69 have the most patients that endure a recurrence of their breast cancer.", # subtitle
    x = "Size of Tumor",  # x axis
    y = "Class of Recurrence",# y axis
    size = "Degree of Malignancy", # size
    frame = "Age", col = "Age") + # frame and column represent age
   transition_states( 
        age, # what are we animating by
        transition_length = 3, #The relative length of the transition.
        state_length = 2) + # The length of the pause between transitions
        ease_aes("linear") + # type of animation
  theme(plot.title = element_text(size = 14, hjust = 0.5, color ="antiquewhite4"), # title size and color
        plot.title.position = "plot", # title adjustments
        plot.subtitle = element_text(size =10,hjust = 0.5, lineheight = 1.1, color = "antiquewhite4"), # subtitle size and color
        axis.title = element_text(size=20,color = "black", hjust = 0.5), # title adjustments
        axis.text = element_text(size = 15,color = "antiquewhite4",hjust = 0.5), # text adjustments
        panel.grid = element_line(linewidth = 0.35, linetype = "dotted", color = "beige"), # background gridlines
        plot.margin = unit(c(1.0, 1.0, 1.0, 1.0), "cm"), # margin adjustments
        plot.background = element_rect(color = "white", fill ="white")) # background color
anim_save(here("Independent Project","Output","bubbleplotclass.gif")) # saving the animation!
```

## **Does age have a huge affect when it comes to severity of malignancy and recurrence probability?**

Through the animation going by age showing the various malignant degrees from 1 being the most acute to 3 being the most severe, those between the ages of 50 to 69 years have more chances of having a recurrent where the cancer will return. Those that are younger still have a chance of a recurrence event but at a lower probability depending on the nature of the tumor and it's malignant rate.


### Bar Plot: Age's Impact on Irradiation Therapy Probability
```{r The Bar Plot, fig.asp= 1.0, fig.align='center', fig.cap= "This animated figure illustrates the age of patients at the time of their diagnosis who have undergone a specific type of radiation therapy known as irradiation therapy which is more invasive and weakening.", echo=FALSE}
ggplot(data = subset(barcancer,!is.na(age)),# drop all na's that were going to be included into the plot
       aes(x = age, # x axis
           fill = irradiat)) + # base this bar plot off of irradiation rate
  geom_bar(position = "dodge") + # what makes it a bar plot
  labs(title = " Rate of Irradiation Therapy\n Based on the Age of the Patients at the Time of Diagnosis",
       subtitle =" As the patients get older, there is a decrease in the amount of individuals that will not undergo \n an extreme form of radiation therapy known as irradiation therapy. ", # detailed title and subtitle 
       x="Patient's Age", # true x axis
       y="Underwent Radiation", # irradiation therapy as y axis
       fill = "Irradiation") + # what the plot will be based off of
  theme_minimal() + # the main theme
  transition_states( 
        age, # what are we animating by
        transition_length = 2, #The relative length of the transition.
        state_length = 1) + # The length of the pause between transitions
        ease_aes("bounce-in-out") + # type of animation
  scale_fill_manual(values = beyonce_palette(2)) + # the color palette all blue :)
  theme(plot.title = element_text(size = 16, hjust = 0.5, color ="darkgoldenrod1"), # title size and color
        plot.title.position = "plot", # title adjustments
        plot.subtitle = element_text(size =10,hjust = 0.5, lineheight = 1.1, color = "darkgoldenrod1"), # subtitle size and color
        axis.title = element_text(size=20,color = "cornflowerblue", hjust = 0.5), # title adjustments
        axis.text = element_text(size = 15,color = "cornflowerblue",hjust = 0.5), # text adjustments
        panel.grid = element_line(linewidth = 0.35, linetype = "dotted", color = "darkgoldenrod1"), # background gridlines
        plot.margin = unit(c(1.0, 1.0, 1.0, 1.0), "cm"), # margin adjustments
        plot.background = element_rect(color = "white", fill ="white")) # background color
        
        anim_save(here("Independent Project","Output","barplotgif.gif")) # save animation!
                   
```
## **Are middle aged individuals more common to undergo irradiation therapy compared to those 50 and older?**

There is no definitive yes or no as shown in the animation, both middle aged and older patients tend to not undergo irradiation therapy due to it's invasive personality of eradicating everything through the use of high energy levels of radiation form various rays such as gamma and x-rays as well as on an atomic level: protons, neutrons, etc and may undergo chemo therapy instead which is the use of drugs such as taxol to destroy and shrink cancerous tumors directly from the source. However, anyone younger than 70-79 had a percentage of individuals undergo irradiation therapy while those who were 70 and above would not have survived to go through with that certain type of therapy due to their age.